Reinforcement-Learning Windturbine Controller

### 2019-11-27 paper/paper.tex
Added Q-Learning and PG section in paper, some DDPG rewrites for action grad clipping

At first, we need to do some assumptions on our environment. The concept environment encapsulates the reality or simulation that the reinforcement learning algorithm works in. This could be an actual cyber-physical system or a simulation. The first assumption to this environment is that this system operates over time and we can discretize time into timesteps. Our environment here is the simulated wind turbine, but it could be a computer game \cite{lillicrapCONTINUOUSCONTROLDEEP2015}, a physics simulation\cite{brockmanOpenAIGym2016} or even an actual existing wind turbine \cite{kolterDesignAnalysisLearning2012}. 
In every timestep, the environment supplies us with a state, also sometimes called observation, and a scalar reward. The observed state doesn't have to represent the entirety of the state of the simulation, in fact it usually is only a small subset. For our wind turbine example, the observed state is limited to what can be measured with sensors on a wind turbine, in an arcade game it could be the pixel output or in a physical simulation it could be some key values in the simulation. 
The reward is a scalar value which judges the current state of the simulation. A higher reward means the current state is better than that of a lower reward. Sometimes, the reward function is obvious from the system, as in an arcade game it would surely be the score or in chess it could be 1 for win, -1 for lose and 0 for not decided yet. However in our paper, we will dedicate a section on how we designed our reward function. In general reinforcement learning theory, it is assumed to exist and be supplied by the environment.
In every timestep, the agent supplies an action in the form of a tensor. This action could be button presses in an arcade game or blade pitch in a windturbine simulation. The mission of the agent is to pick actions which maximize cumulative reward. In other words, it should steer the environment to achieve best performance.
Additionally to the state and reward, an environment can optionally send out a done signal, which indicates a state in which further simulation is not possible and the environment wants to be reset to an initial state. In a chess game, this would be a loss or a win, and with a windturbine that could for example be fatal structural damage.

The algorithms we looked at further assume that the state progressions can be modelled as a Markov-Chain, as in that what happens at timestep t+1 is only dependent on the hidden state at timestep t, not also on for example t-10. As the hidden state could for example be the full RAM of the simulation, we can safely assume that a simulation only calculates the next timestep on what is in the RAM at the current timestep. Also, Markov-Chains allow for several states as an outcome of any state, and that we can model the transition between these state with a probablistic model.

@@

% Bellman equation

\subsection{PG}

@@

@@

As in the beginning we observed our algorithm to get stuck in either \ac{FMM} or \ac{BMM}, we wanted to improve exploration. Exploration is a common problem in reinforcement learning, as algorithms tend to stay in local maxima.

@@

We first added a gaussian noise to our actions and after not seeing any improvemnts, switched to an Ornstein-Uhlenbeck process \cite{uhlenbeckTheoryBrownianMotion1930} as recommended in the DDPG paper \cite[formula (7)]{lillicrapCONTINUOUSCONTROLDEEP2015}. Additionally to the DDPG paper, we let the sigma parameter decay over time, so that the noise gets less when training has proceeded.

@@

The TD3-paper \cite{fujimotoAddressingFunctionApproximation2018} utilizes a random exploration phase, in which a completely random policy \cite{mozerDiscoveringStructureReactive1990} initially explores the environment, before the actual agent switches in. We implemented the random walk at first through gaussian noise and then through an Ornstein-Uhlenbeck process. The later version creates less vibration through less excessive changes in control parameters. As we found that both random policies rarely explore a sensible operational range, we later added an expert PID controller, which implements torque control, and combined that with our Ornstein-Uhlenbeck process. We set the parameters to that controller by experimentation.

@@

As we still didn't see good results, we tried adding parameter-space noise \cite{plappertParameterSpaceNoise2017} to the actor function, as the paper promised better exploration. We could not observe any improvements and deactivated parameter space noise.

\section{Zero Gradients}
All our efforts to aid exploration by adding noise did not yield better exploration, so we had a look at our gradients and found that all our actor gradients are zero. The critic had gradients, but only in the later levels of the net. Eventually, we solved this problem by both simplyfing the architecture and utilizing normalization.

@@

More effectively, we added data normalization. We normalized the data so states, actions and reward always stayed between 1 and -1. This finally fixed our problem of zero gradients and we could observe normal learning. We recommend this technique as a default technique for every reinforcement learning algorithm, as we can not imagine any drawbacks of normalizing input data. As in our simulation, where our state reflects measurements from very different parts of the simulation, some values in the state-array were 8 orders of magnitude different from others. As most of the high-magnitude values are vibrations, the net could not create a link between these and the state-action, while low-magnitude values like rotational speed were not considered due to their small absolute values. At first, we used observed states and rewards from the random exploration phase to calculate maximum/minimum values. As sometimes after random exploration, extreme values were observed, we started normalising on the replay data of a previous run, in which many simulation states were explored.

@@

We observed that our controller delivers high action gradients and reacts strongly between timesteps. This creates vibrations in the structure and especially high blade pitch changes lead to blades breaking or falling off. We tried different methods to circumvent this.

@@

At first, we implemented gradient actionspaces, where the controller output is not an absolute pitch or torque value but a difference to the last output, starting with everything set to 0. This way, we could clip the action gradients by reducing the gradient action space, so that in each timestep, only a fraction of the actual action space could be traversed. If the controller would output a positive change at maximum pitch/torque or a negative change at minimum pitch, this would be ignored. This did effectively limit action gradients, but also stopped us from getting sensible results.

@@

Another idea was that we feed the last n timesteps concatenated to the current observations. We hoped that this way, the agent could derive its own gradients internally. The result of this were high and very noisy q-losses and no convergence whatsoever. Though we still had gradients, we suspect that our relu-activations might have drifted far off to the low end, and thus could not recover.

@@

We eventually solved the problem with the high action gradients by still letting the controller output absolute actions across the entire action spaces, but if gradients exceed a certain threshold, the actual action taken is set to the nearest value with sensible gradients. Though the actions returned from the policy are still wildly unstable, they at least didn't result in death of the turbine anymore and still we didn't hinder convergence through our gradient actionspace. We at first hid this clipping from the replay buffer, but then decided to store the actual values taken, as the models then don't have to also learn the clipping effect. We observed better learning when not hiding the clipping from the replay buffer.

@@

\subsection{Double critic}
Adding twin critic networks such as in \cite{fujimotoAddressingFunctionApproximation2018} eventually solved our problem with loss explosions completely, however the thing didn't learn anymore. To use it in conjunction with \ac{PER}, we needed to decide for a priority formula and decided to take the minimum of the two q values, after the maximum showed biasing towards one of the networks

@@

We name our final variant of \ac{DDPG} \ac{DDDDDPPG}. Our variant is different to \cite{fujimotoAddressingFunctionApproximation2018} in that we don't use target policy smoothing. However, we just have more D.

@@

\section{Comparison to industry controllers}
We aren't even close

@@

We could imagine that first training the policy to output expert actions would improve learning. We imagine a phase at the beginning and after random exploration, where observations from the random exploration phase are fed into the expert controller and the policy is trained to estimate those values directly. This way, after random exploration, the agent would start off with a policy which behaves like the expert controller but can deviate from that in future training, by taking into account q values. A possible algorithm for this could look like:

Alternatively, we could imagine leaving the expert policy training active the whole time, not just once after random exploration, but to reduce the learning rate gradually. This way, in initial training, predicting the expert policies actions would have a higher impact than what the Q-function would suggest, whereas in later training with the smaller learning rate, deviations from the expert policy would be punished less and the policy is allowed to deviate further from the expert policy. This would involve two backward passes per training iteration.

>>>>

At first, we need to do some assumptions on our environment. The concept environment encapsulates the reality or simulation that the reinforcement learning algorithm works in. This could be an actual cyber-physical system or a simulation. The first assumption to this environment is that this system operates over time and we can discretize time into timesteps. Our environment here is the simulated wind turbine, but it could be a computer game \cite{lillicrapCONTINUOUSCONTROLDEEP2015}, a physics simulation\cite{brockmanOpenAIGym2016} or even an actual existing wind turbine \cite{kolterDesignAnalysisLearning2012}. 
In every timestep, the environment supplies us with a state \(x\), also sometimes called observation, and a scalar reward \(r\) for taking an action \(a\) in that state. In theory, the full trajectory \((x_0, a_0, ..., x_t)\) could be needed to describe the full state of the simulation at timestep t \(s_t\). However, the algorithms we looked at assume that the state progressions can be modelled as a Markov-Chain of observations, as in that what happens at timestep t+1 is only dependent on the observation at timestep t, not also on for example t-10. In Markov Chains, there is a state transition probability \(p(x_{t+1}|x_t, a_t)\) for each transition, which we aim to learn. Because we assume the observation \(x\) to be fully descriptive, we will use \(s_t = x_t\). In reality, the observed state doesn't have to represent the entirety of the state of the simulation, in fact it usually is only a small subset. For our wind turbine example, the observed state is limited to what can be measured with sensors on a wind turbine, in an arcade game it could be the pixel output or in a physical simulation it could be some key values in the simulation.

The reward \(r\) is a scalar value which judges the current state of the simulation. A higher reward means the current state is better than that of a lower reward. Sometimes, the reward function is obvious from the system, as in an arcade game it would surely be the score or in chess it could be 1 for win, -1 for lose and 0 for not decided yet. In our paper, we will dedicate a section on how we designed our reward function, whereas in general reinforcement learning theory, it is assumed to exist and be supplied by the environment.

In every timestep, the agent supplies an action \(a\). This action could be button presses in an arcade game or blade pitch in a windturbine simulation. The mission of the agent is to pick actions which maximize cumulative reward. In other words, it should steer the environment to achieve best performance.

Additionally to the state and reward, an environment can optionally send out a done signal \(d\), which indicates a state in which further simulation is not possible and the environment wants to be reset to an initial state. In a chess game, this would be a loss or a win, and with a windturbine that could for example be fatal structural damage.

\subsection{Definitions}

TODO HUGE MESS FOLLOWING
Before diving into the algorithm, we would define some parameters. From our environment, we have
\begin{itemize}
  \item \(s_t\) a single state at timestep t
  \item \(r_t\) the reward at that timestep t
  \item \(a_t\) an action taken at timestep t
  \item \(\varepsilon\) the unknown transition probability matrix of the environment
  \item \(d_t\) a flag (0 or 1) which indicates a terminal condition in the environment
  \item \((s_t, a_t, r_t, d_t, s_{t+1})\) a state transition, going from state \(s_t\), taking action \(a_t\), getting a reward for that action \(r_t\) and a terminal flag \(d_t\) and reaching state \(s_{t+1}\)
\end{itemize}

We furthermore define \(\pi_{prob}: S \rightarrow P(A)\) a probablistic policy that returns a probability distribution over actions for given state and \(\pi: S \rightarrow A\) a deterministic policy that returns a single action for a given state. We will use deterministic policies in all of the paper except for the Q-Learning explanation, which is easier to grasp with probablistic policies.

  \(R_{\pi}(t) = \sum^T_{t'=t}{\gamma^{t'-t}r_{t'}}\) A discounted return, summing up all returns from timestep t up to some point in the far future T while applying a discounting term \(\gamma \in [0, 1]\). The discounting gives less significance for rewards far in the future than rewards close to now. Also, as we need to assume state/action transitions in the future, the return is dependend upon a policy. If we wanted to compute this return, we would need to calculate the expectation over a construct made of all possible trajectories \( (s_0, a_0 \leftarrow \pi_{prob}(s_0), s_1, ...) \), combined with the initial state distribution \(p(s_0)\) and the environment Markov Chain \(p(s_{t}|s_{t-1}, a_{t-1}\). As we don't know the environment probability distributions, we need to find another way.

\begin{itemize}
  \item \((s_0, a_0, s_1, a_1, ... , a_{t-1}, s_t) \rightarrow \tau\) a trajectory of many states and their actions.
  \item \(R(\tau) = \sum_t{r_t} \) an undiscounted return from a trajectory
\end{itemize}

\(V(s)\) is the value of a state defined as accumulating all rewards from this step infinitely into the future: \(V(s) = \sum_t{r_t}\). There are a couple of problems with this function, at first summing infinitely into the future might yield infinite or otherwise unusable results, which is why we introduce a discounting factor \(\gamma\) in \(V(s) = \sum_t{r_t * \gamma^t}\). And furthermore and more obviously, inputting a state to the function, we can not

HUGE MESS OVER

Some terms which we use
\begin{itmeize}
  \item \textbf{Return} A return with respect of a policy and a starting state s gives the accumulative reward of following that policy from state s until the future. As we have a stochastic environment, we have to add an expectation term, so it is the expected accumulative reward of following that policy. Furthermore a return can be discounted or undiscounted. Discounting means we multiply a discounting factor $\gamma \in [0,1]$ to rewards in the fashion of $r_0 \gamma^0 + r_1 \gamma^1 + r_2 \gamma^2 ...$, which results in distant future rewards being weighted less than closer ones.
  \item \textbf{Optimal policy} The theoretical construct of a policy that always takes the best action with respect to accumulative reward. Though we rarely know this policy, there must always be at least one policy which gives the highest returns of all possible policies
  \item \textbf{Value} The value of a state s is the return of that state following the optimal policy.
\end{itemize}

@@

This equation uses the notion of a state-value function, also called Q function. In the Bellman Equation, it describes the total accumulated return until infinity taking an action a in state s and then following the optimal policy afterwards.

The idea behind Q-Learning is that we want to learn the Q function from observations which we have. Assuming we already know the Q value for the next state \(Q(s_{t+1}, a_{t+1})\), we can then calculate the current Q value \( Q(s_t, a_t) = r_t + \gamma Q(s_{t+1}, a_{t+1}) \). Unfortunately, we don't have a definite Q for the next state, so we need to use our estimation which we have already. This estimation is most likely noisy or even plain wrong, so we want to only gradually update our Q. For this, we define our learning target for learning step i as
\begin{equation}
y_i = \mathbb{E}_{s' \sim \varepsilon}[r + \gamma Q(s', \pi(s');\theta_{i-1})]
\end{equation}
For the moment, let's assume the optimal policy $\pi$ which always chooses the action with the highest Q value. Our $y$ is then the current reward added to the highest possible reward in the future. If we had another policy, we would get the future accumulated reward under that policy. We use the parameters $\theta_{i-1}$ from the previous timestep here so that we don't have to derive y when we derive for $\theta_i$, which simplifies the learning term.

Now, on our learning target y, we can define a loss as
\begin{equation}
L(\theta_i) = \mathbb{E}_{s, a}[(y_i - Q(s, a))^2]
\end{equation}
This already looks close to some squared error term. To calculate the expectation value over the environment $\varepsilon$, we will randomly draw samples from the replay buffer and average across them, doing something similar to a Monte-Carlo experiment. The result is equivalent to a normal mean square error calculation.

Deriving for $\theta_i$ and applying the chain rule, we get the loss derivative
\begin{equation}
\Delta_{\theta_i} L(\theta_i) = \mathbb{E}[2(y_i - Q(s, a))\Delta_{\theta_i}Q(s, a)]
\end{equation}
Using our Monte-Carlo like batch from the replay buffer instead of calculating the full expectation term, we end up with stochastic gradient descent.

Thus we just derived a method to train a Q function under the assumption of a policy. In \cite{mnihPlayingAtariDeep2013}, the team had a finite state space and could use the greedy policy $argmax_a Q(s', a)$ by searching the entire action space. We are presented with an infinite action space though, which is why we can't implement a greedy policy without high computational expense. So, for our work, we have to also learn a policy and not just a Q estimator. Luckily, our Q learning rule also works for other policies than the optimal policy, so let's try to derive a way to learn a policy.

\subsection{Policy Gradients}

We will at first give an intuitive explanation of a basic algorithm which forms the basis for many modern reinforcement leaarning algorithms, including the one we chose. The algorithm is called Policy Gradients or REINFORCE and was originally presented by \cite{williamsSimpleStatisticalGradientfollowing}.

Let's assume we have a probablistic policy $\pi: S \rightarrow P(A)$ which is parametrized by $\theta$. An intuitive way of improving this policy could be to increase the gradient of the best possible action $a^*$ for a state $s$:
\begin{equation}
\theta_{i+1} = \theta_i + \alpha \Delta \pi_{\theta_i}(a^*|s)
\end{equation}
We could do this for all states we see and would end up with an iterative way of improving our policy. However, we might not know the best action for any state. Luckily, we derived Q-Learning in the section before and thus have an estimation which action will give which return in this step. We can use this and, for all actions and states, do
\begin{equation}
\theta_{i+1} = \theta_i + \alpha Q(s, a)\Delta \pi_{\theta_i}(a|s)
\end{equation}
This way, actions with higher Q values will receive a higher gradient step, and as Q is constant with respect to $\theta$ we don't have to derive along Q here. However there is still a problem to this. We assume that we use the same $\pi$ for exploration while doing training, so this policy is responsible for taking actions in the environment. This style is also called on-policy learning. So let's assume the policy is poorly initialized at the beginning of our training and gives action 1 a probability 4 times as high as action 2, though taking action 2 would yield a 10\% better return in our Q function. Now, as we are taking action 1 4 times as often as action 2, we will also perform gradient updates on action 1 4 times as often. Though each of the gradient steps is smaller due to the Q factor, the higher frequency of updating action 1 will lead to action 1 getting an advantage over action 2 in this update strategy. To correct for this phenomenon, we could divide the probability of each of the actions from the update term:
\begin{equation}
\theta_{i+1} = \theta_i + \alpha Q(s, a) \frac{ \Delta \pi_{\theta_i}(a|s)}{\pi_{\theta_i}(a|s)}
\end{equation}
We might remember from calculus that $\frac{\Delta x}{x} = \Delta \log x$ so we simplify to
\begin{equation}
\theta_{i+1} = \theta_i + \alpha Q(s, a) \Delta \log \pi_{\theta_i}(a|s)
\end{equation}

This update equation is our basic policy gradient equation.

@@

In neither of these modes, the simulation is within realistic bounds and the observations gotten from the simulation at this part are not credible. After we observed that the controller itself wasn't able to avoid these states on its own, we inhibited reaching these states by outputting a death condition when being in negative rotational areas or high positive. After a death, the simulation gets reset to initial state and the controller has more chances to learn.

@@

As in the beginning we observed our algorithm to get stuck in either \ac{FMM} or \ac{BMM}, we wanted to improve exploration. The term exploration in reinforcement learning means how well the agent explores the state space of the environment. Exploration is a common problem in reinforcement learning, as algorithms tend to stay in local maxima. Intuitively, we are optimizing a policy to move towards an optimal state, while computing what is the optimal state based on what we observed. Thus, if a certain policy already gets locally good values, it will move towards that locally optimal state and never explore out of it, thus the Q network will not know about the good values beyond that.

Additionally, as we discount future rewards, if the slope between the local optimum and the global optimum is too high and wide, the higher return of the global optimum will not propagate until the local optimum because of the discounting done close to it, and though the Q-network knows the global optimum, the gradients along the Q-function still lead into the local optimum.

We suspected the first problem to happen, as neither of the two modes had good rewards, and going out of these modes continously improves rewards. To feed a network with new values closer to an optimum, a common practice in reinforcement learning is to add noise.

@@

In \cite{mnihPlayingAtariDeep2013} they use an epsilon-greedy strategy, where with a certain small probability epsilon, a random action is taken. As we have a continous action space, we can add a little bit of noise towards every action: \(a' = a + n \). We first tried a gaussian noise: \(n \leftarrow \mathcal{N}(\mu, \sigma^2) \). We didn't see any improvements over our old behavior, so we switched to an Ornstein-Uhlenbeck process \cite{uhlenbeckTheoryBrownianMotion1930} as recommended in the \ac{DDPG} paper \cite[formula (7)]{lillicrapCONTINUOUSCONTROLDEEP2015}. Additionally to the \ac{DDPG} paper, we let the sigma parameter decay over time, so that the noise gets less when training has proceeded.

@@

The TD3-paper \cite{fujimotoAddressingFunctionApproximation2018} utilizes a random exploration phase, in which a completely random policy \cite{mozerDiscoveringStructureReactive1990} initially explores the environment, before the actual agent switches in. We implemented the random walk at first through gaussian noise and then through an Ornstein-Uhlenbeck process. The later version creates less vibration through less excessive changes in control parameters. As we found that both random policies rarely explore a sensible operational range, we later added an expert PID controller, which implements torque control, and combined that with our Ornstein-Uhlenbeck process. We set the parameters to that controller by experimentation as we did not need the actual control parameters of the controller. However it could be imagined to use an established controller here. The PID controller was easily able to keep the windturbine in a sensible operation range, and without noise it resulted in a smooth and constant operation. Though our PID controller is certainly inferior to industry-grade controllers, we will view it as an expert controller which is already achieving a certain degree of quality.

@@

As we still didn't see good results, we tried adding parameter-space noise \cite{plappertParameterSpaceNoise2017} to the actor function, as the paper promised better exploration. We could not observe any improvements and thus deactivated parameter space noise for the rest of the experimentation.

\section{Zero Gradients}
All our efforts to aid exploration by adding noise did not yield better results, so we had a look at our gradients and found that all our actor gradients are zero. The critic had gradients, but only in the later levels of the net. The problem of vanishing or exploding gradients is typical for deep learning, but not so common in flatter architectures.

@@

More effectively, we added data normalization. We normalized the data so states, actions and reward always stayed between 1 and -1. This finally fixed our problem of zero gradients and we could observe normal learning. We recommend this technique as a default technique for every reinforcement learning algorithm. As in our simulation, where our state reflects measurements from very different parts of the simulation, some values in the state-array were 8 orders of magnitude different from others. As most of the high-magnitude values are vibrations, the net could not create a link between these and the state-action, while low-magnitude but insightful observations like rotational speed were not considered due to their small absolute values. At first, we used observed states and rewards from the random exploration phase to calculate maximum/minimum values. As sometimes after random exploration, extreme values were observed, we started normalising on the replay data of a previous run. We could also imagine manually setting normalization bounds on low-dimensional problems, as sometimes human knowledge about the possible state space is present.

@@

We observed that our controller delivers high action gradients and reacts strongly between timesteps, sometimes jumping from no pitch/torque all the way to the maximum. This creates vibrations in the structure and especially high blade pitch changes lead to blades breaking or falling off. We tried different methods to circumvent this.

@@

At first, we implemented gradient actionspaces, where the controller output is not an absolute pitch or torque value but a difference to the last output, starting with everything set to 0. For aiding training, we added the actual actions taken to the observation space of the next step. This way, we could clip the action gradients by reducing the gradient action space, so that in each timestep, only a fraction of the actual action space could be traversed. If the controller would output a positive change at maximum pitch/torque or a negative change at minimum pitch, this would be ignored. This did effectively limit action gradients, but also did not lead to any sensible results, so we deactivated this again.

@@

Another idea was that we feed the last n timesteps concatenated to the current observations. We hoped that this way, the agent could derive its own gradients internally. The result of this were high and very noisy q-losses and no sensible results still. Though gradients looked sensible, we suspect that the high Q-losses resulted in too high steps along the gradient and thus diverging behavior.

@@

We eventually solved the problem with the high action gradients by still letting the controller output absolute actions across the entire action spaces, but if gradients exceed a certain threshold, the actual action taken is set to the nearest value with sensible gradients. Though the actions returned from the policy are still wildly unstable, they at least didn't result in death of the turbine anymore and still we didn't hinder convergence through our gradient actionspace. This resulted in our first ever controller keeping the turbine within sensible operational range for at least a part of the training, though diverging quickly after.

We at first hid this clipping from the replay buffer, but then decided to store the actual values taken, as the models then don't have to also learn the clipping effect. We observed better learning when not hiding the clipping from the replay buffer.

\subsection{Pretraining the policy}

We added a small period of direct pretraining of the policy on the actions taken by the expert policy during random exploration. We hoped for a more sensible default policy at the beginning and indeed observed reaching sensible behavior a little earlier.

@@

\subsection{Double critic}
Adding twin critic networks such as in \cite{fujimotoAddressingFunctionApproximation2018} eventually solved our problem with loss explosions completely, however the thing didn't reach sensible states anymore. To use it in conjunction with \ac{PER}, we needed to decide for a priority formula and decided to take the minimum of the two q values, after the maximum showed biasing towards one of the networks.

@@

It must be stated that we did not achieve a single run which stayed within sensible operational range for longer than 30K steps. Even our handcrafted PID controller for which we took the first set of parameters that somehow works widely outperforms our learnt policies, an industry controller is even better. This result in itself is notable, because it needed a whole set of additions to plain \ac{DDPG} to get first sensible results on our windturbine environment. So apparently, our windturbine simulation is more complex to learn than the OpenAI gym examples, which are usually used as benchmarks for \ac{RL} performance. As the final product we developped is close to \cite{fujimotoAddressingFunctionApproximation2018} and has improvements from a wide variety of papers, we conclude that tackling wind turbine control with \ac{RL} will need more research. Though we were not able to outperform industry controllers, the speed in which the area of \ac{RL} is advancing makes us hopeful to achieve good results in the near future. With our work on integrating qblade, we provided an open-source, accessible framework to develop future reinforcement learning controllers for wind turbines.

@@

We tried
\begin{itemize} 
  \item increasing/decreasing both learning rates
  \item increasing/decreasing tau and gamma
  \item increasing/decreasing batch size
  \item increasing/decreasing replay noise
\end{itemize}

\section{Comparison to industry controllers}

@@

We implemented expert policy pretraining. Alternatively, we could imagine leaving the expert policy training active the whole time, not just once after random exploration, and to reduce the learning rate of the expert policy training gradually. This way, in initial training, predicting the expert policies actions would have a higher impact than what the Q-function would suggest, whereas in later training with the smaller learning rate, deviations from the expert policy would be punished less and the policy is allowed to deviate further from the expert policy. This would involve two backward passes per training iteration and would require prediction targets from the expert policy on new seen observations
